---
title: "Summarize and Consolidate Individual Data"
author: "Daniel Hocking"
date: "June 17, 2015"
output: html_document
---

```{r load packages, echo=FALSE, results='hide'}
library(stringr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(scales)
library(devtools)
#devtools::install_github("hadley/tidyr")
library(tidyr)
```

```{r assign directories, echo=FALSE, results='hide'}
# data directory relative to the working directory
data_dir <- "Data/Fish_Data_Processed/Forest_Service_GMNF"

# output directory
out_dir <- "Data/Fish_Data_Processed/Forest_Service_GMNF"

if(!exists(file.path(getwd(), out_dir))) dir.create(file.path(getwd(), out_dir))
```

```{r read data, echo=FALSE, results='hide'}
df_fish2 <- readRDS(file = file.path(data_dir, "GMNF_Individual.RData"))
empty_passes <- readRDS(file = file.path(data_dir, "Empty_Passes.RData"))
```

## Get unique site-visits-pass

Get unique visits and passes across all species to expand to for any single species because not every species or stage captured in each pass on each site-visit

```{r }
df_site_visit_pass <- df_fish2 %>%
  dplyr::group_by(location_id, date, pass) %>%
  dplyr::summarise(count = n())

dim(df_site_visit_pass)

df_max_pass <- df_site_visit_pass %>%
  dplyr::group_by(location_id, date) %>%
  dplyr::summarise(pass_max = max(pass))

empty_passes <- empty_passes %>%
  dplyr::mutate(location_id = paste0(stream, "_", site))



```


## Use Brook Trout data 2001-2013

```{r separate bkt, echo = FALSE, results='hide'}
df_bkt <- df_fish2 %>%
  dplyr::mutate(species = as.character(species),
                year = year(date),
                month = month(date)) %>%
  dplyr::filter(grepl("BKT", species)) %>%
  dplyr::filter(!is.na(date),
                year > 2000)
```

## Assign Stage (YOY/Adult)

Check yearly size distributions by location to assign YOY by year.

```{r, echo = FALSE}
ggplot(group_by(df_bkt, year)) + geom_freqpoly(aes(length, colour = location_id)) + xlim(10, 110) + facet_wrap(~year)

# sites close enough and similar enough that dont' have to do it by stream, just year
ggplot(group_by(df_bkt, year)) + geom_histogram(aes(length), binwidth = 1)  + facet_wrap(~year)

years <- 2001:2013
for(i in 1:length(years)) {
  (g <- ggplot(dplyr::filter(df_bkt, year == years[i] & length >= 50 & length <=150)) + 
     geom_histogram(aes(length), binwidth = 1) + 
     ggtitle(years[i]) + 
     scale_x_continuous(breaks = pretty_breaks(n=20)))
  print(g)
}

yoy_size <- c(83, 80, 85, 81, 88, 87, 87, 90, 85, 84, 85, 90, 85)

df_size <- data.frame(years, yoy_size)
names(df_size) <- c("year", "yoy_length")

# merge and assign stage class
df_bkt <- left_join(df_bkt, df_size, by = "year") %>%
  dplyr::mutate(stage = ifelse(length > yoy_length, "adult", "yoy"),
                stage = ifelse(is.na(length), NA, stage))

# number of fish with no length and no stage
length(df_bkt$length[is.na(df_bkt$length)])
```



```{r}
ggplot(data = df_bkt, aes(location_id, year)) + geom_point()

count_table <- addmargins(with(df_bkt, table(location_id, year)))

pa_table <- with(df_bkt, table(location_id, year))
pa_table[pa_table>0] <- 1
rbind(pa_table, colSums(pa_table))
```

## determine the number of bkt per pass - need info on whether any fish captured to get 0s

# summarize
```{r summarise}
df_sum <- df_bkt %>%
  dplyr::group_by(location_id, date, pass, species, stage) %>% # need to classify YOY
  dplyr::summarise(count = n(),
                   width_ft = mean(width),
                   area_ft2 = mean(area))

df_sum <- dplyr::arrange(ungroup(df_sum), location_id, date, stage, pass)
head(df_sum, 20)
summary(df_sum)
dim(df_sum)

# how to handle NA in length/stage?
dplyr::filter(df_bkt, is.na(length))

# it seems like NA represent zero captures for that species on that pass. I can throw these out then use tidyr::complete (expand.grid) to set these as 0 for all life age/stages.

df_sum <- df_sum %>%
  dplyr::filter(!is.na(stage)) %>%
  dplyr::mutate(visit = paste0(location_id, "_", as.character(date)))

# need to expand to all passes for all loction-dates-species
# use new tidyr wrapper for expand.grid
df_sum2 <- tidyr::complete(df_sum, c(visit), pass, stage, fill = list(count = 0)) %>%
  dplyr::arrange(location_id, date, species, stage, pass) %>%
  dplyr::mutate(year = year(date),
                month = month(date))
head(df_sum2, 20)
dim(df_sum2)
summary(df_sum2)

# check that the size I want
n_site_visits <- length(unique(df_sum2$visit))
n_pass <- 3
n_stage <- 2
n_species <- 1

foo <- expand.grid(site_visits = unique(df_sum2$visit), pass = 1:3, stage = c("adult", "yoy"), species = "BKT")

# assume ALWAYS have 3 passes per visit
n_rows_expect <- n_site_visits * n_pass * n_stage * n_species

n_rows_expect == dim(df_sum2)[1]

df_sum2 <- dplyr::arrange(df_sum2, location_id, date, stage, pass)
head(df_sum2, 20)
tail(df_sum2, 20)
print(df_sum2[100:119, ])

```

## block nets used pre-2011

block nets were no longer used starting in 2011 because of issues with disease/invasives

```{r block nets, echo=FALSE}
df_nets <- data.frame(year = as.integer(unique(df_sum2$year)))

df_nets <- na.omit(df_nets)

df_nets <- df_nets %>%
  dplyr::mutate(net = ifelse(year < 2011, TRUE, FALSE))

df_sum2 <- left_join(df_sum2, df_nets, by = c("year"))
str(df_sum2)
summary(df_sum2)
```


## Export all GMNF block net count data

```{r GMNF net counts}
saveRDS(df_sum2, file = file.path(out_dir, "GMNF_Blocknet_Counts.Rdata"))
```


## extract white river data for spatial analysis

```{r white river}
load(file.path(out_dir, "White_River_Network.RData"))

locations_wr <- dplyr::filter(family, !grepl("N_", child_b))$child_b

pa_table <- with(dplyr::filter(df_bkt, location_id %in% locations_wr), table(location_id, year))
pa_table[pa_table>0] <- 1
pa_table <- rbind(pa_table, colSums(pa_table))
cbind(pa_table, rowSums(pa_table))

df_grf <- df_sum2 %>%
  dplyr::filter(location_id %in% locations_wr) %>%
  dplyr::filter(net == TRUE) %>%
  dplyr::mutate(length_sample = 12*2.54/100*(area_ft2/width_ft),
                width = 12*2.54/100*width_ft) %>%
  dplyr::select(-area_ft2, -width_ft)

df_grf <- dplyr::arrange(df_grf, location_id, date, stage, pass)
head(df_grf, 20)

# need to expand to all passes for all loction-dates-species
# use new tidyr wrapper for expand.grid
df_grf <- tidyr::complete(df_grf, c(visit), pass, stage, fill = list(count = 0)) %>%
  dplyr::arrange(location_id, date, species, stage, pass)

# check that the size I want
n_site_visits <- length(unique(df_grf$visit))
n_pass <- 3
n_stage <- 2
n_species <- 1

# assume ALWAYS have 3 passes per visit
n_rows_expect <- n_site_visits * n_pass * n_stage * n_species

n_rows_expect == dim(df_grf)[1]

saveRDS(df_grf, file = file.path(out_dir, "White_River_Trout.Rdata"))
```


       
       
       
       
       
       
       
       
